---
title: "Stats"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

# Statistical Inferences

```{r}
# setup
library(gapminder)
gap <- gapminder
```

## Statistical Distributions

Since R was developed by statisticians, it handles distributions and simulation seamlessly.

All commonly-used distributions have functions in R. Each distribution has a family of functions: 

* `d` - Probability density/mass function, e.g., `dnorm()`
* `r` - Generate a random value, e.g., `rnorm()`
* `p` - Cumulative distribution function, e.g., `pnorm()`
* `q` - Quantile function (inverse CDF), e.g., `qnorm()`

Let's see some of these functions in action with the normal distribution (mean 0, standard deviation 1):

```{r, fig.cap = ""}
dnorm(1.96) # Probability density of 1.96 from the normal distribution
rnorm(1:10) # Get 10 random values from the normal distribution
pnorm(1.96) # Cumulative distribution function
qnorm(.975) # Inverse cumulative distribution function
```

We can also use these functions on other distributions:

* `rnorm()` # Normal distribution
* `runif()` # Uniform distribution
* `rbinom()` # Binomial distribution
* `rpois()` # Poisson distribution
* `rbeta()` # Beta distribution
* `rgamma()` # Gamma distribution
* `rt()` # Student t distribution
* `rchisq()`. # Chi-squared distribution

```{r}
rbinom(0:10, size = 10, prob = 0.3)
dt(5, df = 1)

x <- seq(-5, 5, length = 100)
plot(x, dnorm(x), type = 'l')
```

### Sampling and Simulation

We can draw a sample with or without replacement with `sample`.

```{r}
sample(1:nrow(gap), 20, replace = FALSE)
```

`dplyr` has a helpful `select_n` function that samples rows of a dataframe.

```{r}
small <- sample_n(gap, 20)
nrow(small)
```

Here's an example of some code that would be part of a bootstrap:

```{r}
gap <- read.csv("data/gapminder-FiveYearData.csv", stringsAsFactors = F)

# Actual mean
mean(gap$lifeExp, na.rm = TRUE)

# Here's a bootstrap sample:
smp <- sample_n(gap, size = nrow(gap), replace = TRUE) 
mean(smp$lifeExp, na.rm = TRUE)
```

### Random Seeds

A few key facts about generating random numbers:

* Random numbers on a computer are *pseudo-random*; they are generated deterministically from a very, very, very long sequence that repeats.
* The *seed* determines where you are in that sequence.

To replicate any work involving random numbers, make sure to set the seed first. The seed can be arbitrary -- pick your favorite number.

```{r}
set.seed(1)
vals <- sample(1:nrow(gap), 10)
vals

vals <- sample(1:nrow(gap), 10)
vals

set.seed(1)
vals <- sample(1:nrow(gap), 10)
vals
```

### Challenges

#### Challenge 1. {-}

Generate 100 random Poisson values with a population mean of 5. How close is the mean of those 100 values to the value of 5? 

#### Challenge 2. {-}

What is the 95th percentile of a chi-square distribution with 1 degree of freedom? 

#### Challenge 3. {-}

What is the probability of getting a value greater than 5 if you draw from a standard normal distribution? What about a t distribution with 1 degree of freedom?
