---
title: "Strings"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
always_allow_html: true
---

# Strings and Regular Expressions

This unit focuses on character (or "string") data. We will explore

1. **String basics**, like concatinating and subsettings.
2. **Regular expressions**, a powerful cross-language tool for working with string data.
3. **Common tools** that take regex and apply them to real problems.

This chapter will focus on the __`stringr`__ package for string manipulation. `stringr` is not part of the core `tidyverse` because you do not always have textual data, so we need to load it explicitly.

```{r message = FALSE}
library(tidyverse)
library(stringr)
```

## String Basics

This unit focuses on character (or "string") data. We will focus on **string basics**, such as concatinating and subsettings.

### Creating Strings

You can create strings with either single quotes or double quotes. Unlike other languages, there is no difference in behavior. I recommend always using `"`, unless you want to create a string that contains multiple `"`.

```{r}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

### Escape and Special Characters 

Single and double quotes are known as "metacharacters," meaning that they have special meaning to the R language. To include a literal single or double quote in a string you can use `\` to "escape" it:

```{r}
double_quote <- "\"" # or '"'
single_quote <- '\'' # or "'"
```

That means if you want to include a literal backslash, you will need to double it up: `"\\"`.

Beware that the printed representation of a string is not the same as the string itself, because the printed representation shows the escapes. To see the raw contents of the string, use `writeLines()`:

```{r}
x <- c("\"", "\\")
x
writeLines(x)
```

There are a handful of other special characters. The most common are `"\n"`, newline, and `"\t"`, tab, but you can see the complete list by requesting help on `"`: `?'"'`, or `?"'"`. 

Sometimes you will also see strings like `"\u00b5"`. This is a way of writing non-English characters that works on all platforms:

```{r}
x <- "\u00b5"
x
```

Multiple strings are often stored in a character vector, which you can create with `c()`:

```{r}
c("one", "two", "three")
```

### String Length

Base R contains many functions to work with strings, but we will avoid them because they can be inconsistent, which makes them hard to remember. 

Instead, we will use functions from `stringr`. `stringr` contains functions with more intuitive names, and all start with `str_`. For example, `str_length()` tells you the number of characters in a string:

```{r}
str_length(c("a", "R for data science", NA))
```

The common `str_` prefix is particularly useful if you use RStudio, because typing `str_` will trigger autocomplete, allowing you to see all `stringr` functions:

### Combining Strings

To combine two or more strings, use `str_c()`:

```{r}
str_c("x", "y")
str_c("x", "y", "z")
```

Use the `sep` argument to control how they are separated:

```{r}
str_c("x", "y", sep = ", ")
```

`str_c()` is vectorised, and it automatically recycles shorter vectors to the same length as the longest:

```{r}
x <- c("a", "b", "c")
str_c("prefix-", x)
```

To collapse a vector of strings into a single string, use `collapse`:

```{r}
x <- c("x", "y", "z")
str_c(x, sep = ", ") # This will not work
str_c(x, collapse = ", ") # But this will
```

### Subsetting Strings

You can extract parts of a string using `str_sub()`. As well as the string, `str_sub()` takes `start` and `end` arguments, which give the (inclusive) position of the substring:

```{r}
x <- c("Rochelle is the GOAT")
str_sub(x, 1, 8)

# Negative numbers count backwards from the end
str_sub(x, -8, -1)
```

You can also use the assignment form of `str_sub()` to modify strings:

```{r}
x <- c("Rochelle is the GOAT")
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x
```

### Locales

Above I used `str_to_lower()` to change the text to lower case. You can also use `str_to_upper()` or `str_to_title()`. However, changing case is more complicated than it might at first appear, because different languages have different rules for changing case. You can pick which set of rules to use by specifying a locale:

```{r}
# Turkish has two i's (with and without a dot), and it
# has a different rule for capitalising each of them:
str_to_upper(c("i", "ı"))
str_to_upper(c("i", "ı"), locale = "tr")
```

The locale is specified as a ISO 639 language code, which is a two- or three-letter abbreviation. If you do not already know the code for your language, [Wikipedia](https://en.wikipedia.org/wiki/List_of_ISO_639-1_codes) has a good list. If you leave the locale blank, it will use the current locale, as provided by your operating system.

Another important operation that is affected by the locale is sorting. The base R `order()` and `sort()` functions sort strings using the current locale. If you want robust behavior across different computers, you may want to use `str_sort()` and `str_order()`, which take an additional `locale` argument:

```{r}
x <- c("apple", "eggplant", "banana")
str_sort(x, locale = "en")  # English
str_sort(x, locale = "haw") # Hawaiian
```

### Challenges

#### Challenge 1. {-}

In code that does not use `stringr`, you will often see `paste()` and `paste0()`. What is the difference between the two functions? What `stringr` function are they equivalent to? How do the functions differ in their handling of `NA`?
    
#### Challenge 2. {-}

In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

#### Challenge 3. {-}

Use `str_length()` and `str_sub()` to extract the middle character from 
    a string. What will you do if the string has an even number of characters?

#### Challenge 4. {-}

What does `str_trim()` do? What is the opposite of `str_trim()`?



