---
title: "Basic Syntax"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Boolean Expressions 

Boolean expressions are logical statements that are either true or false. For our purposes, we will often use Boolean expressions to compare quantities. For instance, the Boolean expression `1 < 2` is true, whereas the Boolean expression `1 > 2` is false.

When you type a Boolean expression in R, R will output `TRUE` if the expression is true and `FALSE` if the expression is false. 

### Boolean Operators

| Operator  | Description              |
|-----------|--------------------------|
| `<`         | less than                |
| `<=`        | less than or equal to    |
| `>`         | greater than             |
| `>=`        | greater than or equal to |
| `==`        | exactly equal to         |
| `!=`        | not equal to             |
| `%in%`      | is an item of a set      |


Note that we use a double equal sign `==` to check whether two values are equal. Typing `a = b` would set the value of `a` equal to the value of `b`.

Here are some examples of Boolean expressions in action:

```{r}
1 < 2
1 > 2
1 == 2

value_1 <- 1
value_1 > 0

value_2 <- value_1 + 10
value_1 + value_2 <= 12
```

### Logical Operators

In practice, we often use two or more conditions to decide what to do. To combine different conditions, there are three logical operators:

| Operator  | Description              |
|-----------|--------------------------|
| `x & y`     | `x` AND y              |
| `x | y`     | `x` OR y               |
| `!x`        | Not `x`                |

First, `&` is similar to **AND** in English, that is, `x & y` is true only if both `x` and `y` are true. Second,  is similar to **OR** in English, that is, A || B is false only if both `x` and `y` are false. Third, `!` is similar to **NOT** in English, that is, `!x` is true only if `x` is false.

```{r}
x <- 10
y <- 20

# check to see if values are between 5 and 15
x > 5 & x < 15
y > 5 & y < 15

# more complex chains
(x > 5 & x < 15) & (x > y & y < 15)
(x > 5 & x < 15) | (x > y & y < 15)
(x > 5 & x < 15) & !(x > y & y < 15)
```

Different operator take different precendence. It is always a good practice to use brackets to control operation ordering.

### Boolean Vectors

The nice thing about R is that you can use these comparison operators also on vectors. As with many expressions in R, the Boolean expressions discussed above are all *vectorized.* We will learn more about vectors and vectorization later in this class, but here is a quick example:

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x > 3
```

This command tests for every element of the vector if the condition stated by the comparison operator is `TRUE` or `FALSE`.

### Boolean Vectors in Action

Boolean vectors are partially what makes R so magical. Check out the example below and examine each line. We will cover subsetting operations later, but pay attention to the work of boolean expressions and logical operators.

```{r}
# An example
x <- c(1:10)
x[(x>8) | (x<5)]

# How it works
x <- c(1:10)
x
x > 8
x < 5
x > 8 | x < 5
x[c(T,T,T,T,F,F,F,F,T,T)]
```

### Challenges

#### Challenge 1. {-}

In the partially written code below, `vector_1` and `vector_2` each contain 10 values. Using the fact that Boolean expressions are vectorized, write code that outputs:

1. A vector of length 10 such that element `i` of this vector equals `TRUE` if `vector_1[i]` is less than `vector_2[i]` and equals `FALSE` otherwise.
2. The number of times that element `i` of `vector_1` is less than element `i` of `vector_2`, using the `sum` function.

```{r}
vector_1 <- c(1, 2, 4, 5, 3, 7, 8, 7, 1, 2)
vector_2 <- c(1, 3, 4, 4, 5, 10, 6, 8, 9, 1)

# YOUR CODE HERE
```

