---
title: "03_Challenge-Solutions"
output: html_document
---

## 1_R-Variables Challenges

#### Challenge 1. {-}

Make 3 variables: name (with your full name), city (where you were born), and year (when you were born).

A:
```{r}

name <- "Jenna Gibson"

city <- "Stillwater"

year <- 1989

```

#### Challenge 2. {-}

Draw a table showing the values of the variables in this program after each statement is executed.

In simple terms, what do the last three lines of this program do?

```{r}

lowest <- 1.0

highest <- 3.0

temp <- lowest

lowest <- highest

highest <- temp

```

A: 
Values:
1) lowest = 1
2) highest = 3
3) temp = 1
4) lowest = 3
5) highest = 1

The last three lines overwrite existing variables, changing their value.


#### Challenge 3. {-}

What is the final value of `position` in the program below? Try to predict the value without running the program, then check your prediction.

```{r}

initial <- "left"
position <- initial
initial <- "right"

```

A:
Position should have the value of "left."

#### Challenge 4. {-}

Why does the following code fail?

```{r eval = F}

age == 31

```

And the following?

```{r eval = F}
31 <-  age
```


A:
We do not have a variable called "age," so R cannot evaluate whether the variable age has the value of 31 or not.

Numbers cannot be variable names, and also we still do not have a variable called age that we can assign to things.


## 2_R-Functions Challenges

#### Challenge 1. {-}

Explain, in simple terms, the order of operations in the following program: when does the addition happen, when does the subtraction happen, when is each function called, etc.
What is the final value of `radiance`?

```{r}

radiance <- 1.0
radiance <- max(2.1, 2.0 + min(radiance, 1.1 * radiance - 0.5))

```

A:
1) Assigning the value of 1 to the variable radiance.
2) 1.1 * radiance = 1.1
3) 1.1 - 0.5 = 0.6
4) Smaller number between radiance (1.0) and 0.6 = 0.6
5) 2.0 + 0.6 = 2.6
6) Larger number between 2.1 and 2.6 = 2.6

The final value of radiance is 2.6.

#### Challenge 2. {-}

Run the following code and, using the help files for `max`, explain why it returns the result it does:

```{r eval = F}

rich <- "gold"
poor <- "tin"
max(rich, poor)

?max
```

A:
The max function takes either numeric or character inputs. For character inputs, it will evaluate based on alphabetical order. Max returns the value that is last in alphabetical order --- in this case, "tin."


## 3_R-Syntax Challenges

#### Challenge 1. {-}

1. Make a variable `year` and assign it as the year you were born.
2. Coerce that variable to a string and assign it to a new variable `year_string`.
3. Someone in your class says they were born in 2001. Really? Really. Find out what your age difference is, using only `year_string`.

```{r}

year <- 1989

year_string <- as.character(year)

2001 - as.numeric(year_string)

```


#### Challenge 2. {-}

Change the following code to make the output `TRUE`:

```{r}
#Original Code
val_1 <- F
val_2 <- "F"
class(val_1) == class(val_2)

#A: Fixed Code
#Can just add " " to val_1 to make it a character vector originally, or:

val_1 <- F
val_2 <- "F"
class(val_1) == class(as.logical(val_2))

```





## 4_R-Syntax Challenges

#### Challenge 1. {-}

In the partially written code below, `vector_1` and `vector_2` each contain 10 values. Using the fact that Boolean expressions are vectorized, write code that outputs:

1. a vector of length 10 such that element `i` of this vector equals `TRUE` if `vector_1[i]` is less than `vector_2[i]` and equals `FALSE` otherwise.
2. the number of times that element `i` of `vector_1` is less than element `i` of `vector_2`, using the `sum` function.

```{r}
vector_1 <- c(1, 2, 4, 5, 3, 7, 8, 7, 1, 2)
vector_2 <- c(1, 3, 4, 4, 5, 10, 6, 8, 9, 1)

vector_1 < vector_2
sum(vector_1 < vector_2)
```

