---
title: "Preprocessing"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Sentiment Analysis and Dictionary Methods

To demonstrate sentiment analysis, we are going to explore lyrics from Taylor Swift songs. 

Road the code below to get started:

```{r message = F}
require(tm)
require(tidytext)
require(tidyverse)
require(stringr)
require(textdata)
```

### Preprocessing and Setup

First, we must preprocess the corpus. Create a document-term matrix from the `lyrics` column of the `ts` dataframe. Complete the following preprocessing steps:

- Convert to lower.
- Remove stop words.
- Remove numbers.
- Remove punctuation.

**Think**: Why is stemming inappropriate for this application?

```{r}
ts <- read.csv("data/taylor_swift.csv")

# Preprocess and create DTM
docs <- Corpus(VectorSource(ts$lyrics))

dtm <- DocumentTermMatrix(docs,
           control = list(tolower = TRUE,
                          removeNumbers = TRUE,
                          removePunctuation = TRUE,
                          stopwords = TRUE
                         ))

# Convert to dataframe
dtm <- as.data.frame(as.matrix(dtm))
```

#### Sentiment Dictionaries {-}

We are going to use sentiment dictionaries from the `tidytext` package. Using the `get_sentiments` function, load the "bing" dictionary and store it in an object called `sent`. 

```{r}
sent <- get_sentiments("bing")
head(sent)
```

We will now add a column to `sent` called `score`. This column should hold a "1" for positive words and "-1" for negative words.

```{r}
sent$score <- ifelse(sent$sentiment=="positive", 1, -1)
```

### Scoring the Songs

We are now ready to score each song. 

(**NB**: There are probably many ways to program a script that performs this task. If you can think of a more elegant way, go for it!)

First, we will create a dataframe that holds all the words in our DTM along with their sentiment score.

```{r}
# Get all the words in our DTM and put them in a dataframe
words = data.frame(word = colnames(dtm), stringsAsFactors = F)
head(words)

# Get their sentiment scores
words_sent <- words %>%
  left_join(sent) %>% 
  mutate(score = replace_na(score, 0))
```

We can now use matrix algebra (!!) to multiply our DTM by the scoring vector. This will return to us a score for each document (i.e., song).

```{r}
# Calculate documents scores with matrix algebra! 
doc_scores <- as.matrix(dtm) %*% words_sent$score

# Put the scores in the original documents dataframe
ts$sentiment <- doc_scores
```

Which song is happiest? Go listen to the song and see if you agree.

### Challenges

#### Challenge 1. {-} 

Using the code we wrote above, make a function that accepts 1) a vector of texts and 2) a sentiment dictionary (i.e., a dataframe with words and scores) and returns a vector of sentiment scores for each text.

```{r eval = F}
sentiment_score <- function(texts, sent_dict){
  
  # YOUR CODE HERE

  return(doc_scores)
}
  
# Uncomment to test it out!
# sentiment_score(ts$lyrics, sent_dict)
```

#### Challenge 2. {-}

Using the function you wrote above, find out what the most and least positive Taylor Swift album is.

```{r eval = F}
# YOUR CODE HERE
```

