---
title: "08-99_Challenge_Solutions"
output: html_document
---

```{r}
library(ggplot2)
library(gapminder)

gap <- gapminder
```

## 8-2_ggplot Challenges

#### Challenge 1. {-}



## 8-2_ggplot Challenges

#### Challenge 1. {-}

Modify the example so that the figure visualizes how life expectancy has changed over time:

Hint: The `gapminder` dataset has a column called `year`, which should appear on the x-axis.

```{r}
# Original
ggplot(data = gap, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point()

# Challenge
ggplot(data = gap, aes(x = year, y = lifeExp)) + 
  geom_point()
```

#### Challenge 2. {-}

Switch the order of the point and line layers from the previous example. What happened?

```{r}
# Original
ggplot(data = gap, aes(x=year, y=lifeExp, by=country)) + 
  geom_line(aes(color=continent)) + 
  geom_point()

# Challenge
ggplot(data = gap, aes(x=year, y=lifeExp, by=country)) + 
  geom_point() + 
  geom_line(aes(color=continent)) 
```

Answer: The point layer now appears underneath the line layer.

#### Challenge 3. {-}

Modify the color and size of the points on the point layer in the previous example so that they are fixed (i.e., not reflective of continent).

Hint: Do not use the `aes` function.

```{r}
# Original
ggplot(data = gap, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(aes(color=continent)) + 
  scale_x_log10() + 
  geom_smooth(method="lm", size = 1.5)

# Challenge
ggplot(data = gap, aes(x = gdpPercap, y = lifeExp)) + 
  geom_point(color = "red", size = 1) + 
  scale_x_log10() + 
  geom_smooth(method="lm", size = 1.5)
```

#### Challenge 4. {-}

Create a density plot of GDP per capita, filled by continent.

Advanced: 
 * Transform the x-axis to better visualize the data spread. 
 * Add a facet layer to panel the density plots by year.

```{r}
# Create a density plot of GDP per capita, filled by continent.
ggplot(gap, aes(x = gdpPercap)) +
  geom_density(aes(color = continent))

# Transform the x-axis to better visualise the data spread. 
ggplot(gap, aes(x = gdpPercap)) +
  geom_density(aes(color = continent)) +
  scale_x_log10()

# Add a facet layer to panel the density plots by year.
ggplot(data = gap, aes(x = gdpPercap)) + 
  geom_density() + 
  scale_x_log10() +
  facet_wrap( ~ year)
```

