---
title: "Indexing Vectors"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Subsetting Vectors

Sometimes we want to isolate elements of a vector for inspection, modification, etc. This is often called **indexing** or **subsetting**.

> By the way, indexing begins at 1 in R, unlike many other languages that index from 0.

### Subsetting Types

Let's explore the different types of subsetting with a simple vector, `x`:

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
```

Note that the number after the decimal point gives the original position in the vector.

There are four things you can use to subset a vector: 

#### 1.   __Positive integers__ return elements at the specified positions. {-}

The simplest way to subset a vector is with a single integer:

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x[1]
x[3]
```

We can also index multiple values by passing a vector of integers:

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x[c(3, 1)]
```

Note that you *have* to use `c` inside the `[` for this to work!

More examples:

```{r}
# `order(x)` gives the index positions of smallest to largest values
(x <- c(2.1, 4.2, 3.3, 5.4))
order(x)

# Use this to order values
x[order(x)]
x[c(1, 3, 2, 4)]
```

#### 2.   __Negative integers__ omit elements at the specified positions. {-}

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x[-1]
x[-c(1, 3)]
```

You cannot mix positive and negative integers in a single subset:

```{r, error = TRUE}
x <- c(2.1, 4.2, 3.3, 5.4)
x[c(-1, 2)]
```

#### 3. __Character vectors__ return elements with matching names. This only works if the vector is named. {-}
 
```{r}
x <- c(2.1, 4.2, 3.3, 5.4)

# Apply names
names(x) <- c("a", "b", "c", "d")

# Subset using names
x["d"]
x[c("d", "c", "a")]
```

#### 4.  __Logical vectors__ select elements where the corresponding logical value is `TRUE`. {-}

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x[c(TRUE, TRUE, FALSE, FALSE)]
```

Logical subsetting is the most useful type of subsetting, because you use it to subset based on __comparative__ statements.

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x > 3
```

This command tests if the condition stated by the comparison operator is `TRUE` or `FALSE` for every element of the vector, and it returns a logical vector! 

We can now pass this statement between the square brackets that follow `x` to subset only those items that match `TRUE`:

```{r}
x <- c(2.1, 4.2, 3.3, 5.4)
x[x > 3]

# With !
!x > 5 
x[!x > 5]

# With %in% 
x %in% c(3.3, 4.2)
x[x %in% c(3.3, 4.2)]
```

#### Challenge. {-}

Subset `country_vector` below to return every value EXCEPT `"Canada"` and `"Brazil"`.

```{r}
country_vector<-c("Afghanistan", "Canada", "Sierra Leone", "Denmark", "Japan", "Brazil")

# Do it using positive integers.

# Do it using negative integers.

# Do it using a logical vector.

# Do it using a conditional statement (and an implicit logical vector).
```
