---
title: "Indexing Lists"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Subsetting Lists

Subsetting a list works in the same way as subsetting an atomic vector. However, there is one important difference: `[` will always return a list. `[[` and `$`, as described below, let you pull out the components of the list.

The ["pepper shaker photos"](https://r4ds.had.co.nz/vectors.html#lists-of-condiments) in R for Data Science are a splendid visual explanation of the different ways to get stuff out of a list. Highly recommended.

Let's illustrate with the following list `my_list`:

```{r}
my_list <- list(a = 1:3, 
                b = "a string", 
                c = pi, 
                d = list(-1, -5))
```

### With `[`

`[` extracts a sub-list where the result will always be a list. Like with vectors, you can subset with a logical, integer, or character vector.

```{r}
my_list[1]
str(my_list[1])

my_list[1:2]
str(my_list[1:2])
```

### With `[[`

`[[` extracts a single _component_ from a list. In other words, it removes that hierarchy and returns whatever object is stored inside.

```{r}
my_list[[1]]
str(my_list[[1]])

# Compare to
my_list[1]
str(my_list[1])
```

The distinction between `[` and `[[` is really important for lists, because `[[` drills down into the list while `[` returns a new, smaller list. 

>  "If list `x` is a train carrying objects, then `x[[5]]` is
> the object in car 5; `x[4:6]` is a train of cars 4-6." 
>
> --- @RLangTip


### with `$`

`$` is a shorthand for extracting a single named element of a list. It works especially well when coupled with tab completion. 

```{r}
my_list$a
```

### Challenges

#### Challenge 1. {-}

What are the four basic types of atomic vectors? How does a list differ from an atomic vector?

#### Challenge 2. {-}

Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?

#### Challenge 3. {-}

Create three vectors and combine them into a list. Assign them names.

#### Challenge 4.  {-}

If `x` is a list, what is the class of `x[1]`?  How about `x[[1]]`?

#### Challenge 5.  {-}

Take a look at the linear model below:

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
summary(mod)
```

Extract the R squared from the model summary.
