---
title: "Lists"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Lists

Lists are different from vectors because their elements can be of __any type__. Lists are sometimes called recursive vectors, because a list can contain other lists. This makes them fundamentally different from vectors.

In data analysis, you will not make lists very often, at least not consciously, but you should still know about them. Why?

1. Dataframes are lists! They are a special case where each element is an atomic vector, all having the same length.
2. Many functions will return lists to you, and you will want to extract goodies from them, such as the p-value for a hypothesis test or the estimated error variance in a regression model.

### Creating Lists

You construct lists by using `list()` instead of `c()`: 

```{r}
x <- list(1, "a", TRUE, c(4, 5, 6))
x
```

### Naming Lists

As with vectors, we can attach names to each element on our list:

```{r eval = F}
my_list <- list(name1 = elem1, 
                name2 = elem2)
```

This creates a list with components that are named `name1`, `name2`, and so on. If you want to name your lists after you have created them, you can use the `names()` function as you did with vectors. The following commands are fully equivalent to the assignment above:

```{r eval = F}
my_list <- list(elem1, elem2)
names(my_list) <- c("name1", "name2")
```

### List Structure

A very useful tool for working with lists is `str()`, because it focuses on reviewing the structure of a list, not its contents.

```{r}
x <- list(a = c(1, 2, 3),
          b = c("Hello", "there"),
          c = 1:10)
str(x)
```

A list does not print to the console like a vector. Instead, each element of the list starts on a new line.

```{r}
x_vec <- c(1,2,3)
x_list <- list(1,2,3)
x_vec
x_list
```

Lists are used to build up many of the more complicated data structures in R. For example, both dataframes and linear model objects (as produced by `lm()`) are lists:

```{r}
head(mtcars)
is.list(mtcars)
mod <- lm(mpg ~ wt, data = mtcars)
is.list(mod)
```

You could say that a list is some kind of super data type: You can store practically any piece of information in it!

For this reason, lists are extremely useful inside functions. You can "staple" together lots of different kinds of results into a single object that a function can return.

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
str(mod)
```
