---
title: "Matrices"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Matrices

Matrices are like 2-d vectors, that is, they are a collection of elements of the same data type (numeric, character, or logical) arranged into a fixed number of rows and columns.

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```

> General arrays are available in R, where a matrix is an important special case having dimension 2.

Practically speaking, matrices are good for large tables of numbers. However, as social scientists, we rarely work with purely numerical data. 

By definition, if you want to combine different types of data (one column numbers, another column characters...), you want a **dataframe**, not a matrix. 

Let’s make a simple matrix and give it decent row and column names. You will see familiar or self-explanatory functions below for getting to know a matrix.

```{r}
## Do not worry if the construction of this matrix confuses you; 
## just focus on the product
m <- outer(as.character(1:4), as.character(1:4),
              function(x, y) {
                paste0('x', x, '-', y)
                })
m
str(m)
class(m)
dim(m)
nrow(m)
ncol(m)
rownames(m)

rownames(m) <- c("row1", "row2", "row3", "row4")
colnames(m) <- c("col1", "col2", "col3", "col4")

m
```

## Indexing a Matrix

Similar to vectors, you can use the square brackets `[ ]` to select one or multiple elements from a matrix. But whereas vectors have one dimension, matrices have two dimensions. We therefore have to use two subsetting vectors – one for rows and another for columns to select – separated by a comma. Blank subsetting is also useful because it lets you keep all rows or all columns.

```{r}
m[2, 3] # Selects the value at the second row and third column

m[2, ] # We get row 2

m[ , 3, drop = FALSE] # We get column 3

dim(m[ , 3, drop = FALSE]) # We get column 3 as a 4 x 1 matrix

m[c("row1", "row4"), c("col2", "col3")] # We get rows 1, 4 and columns 2, 3

m[-c(2, 3), c(TRUE, TRUE, FALSE, FALSE)] # Wacky but possible
```
