---
title: "Dataframes"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Dataframes

Dataframes are a very important data type in R. It is pretty much the *de facto* data structure for most tabular data and it is also what we use for statistics. 

Hopefully the slog through vectors, matrices, and lists will be redeemed by greater prowess at manipulating `data.frames`. Why should this be true?

1. A dataframe is a *list*.
2. The list elements are the variables, and they are atomic vectors.
3. Dataframes are rectangular, like their matrix friends, so your intuition – and even some syntax – can be borrowed from the matrix world.

**NB**: You might have heard of "tibbles," used in the `tidyverse` suite of packages. Tibbles are like dataframes 2.0, tweaking some of the behavior of dataframes to make life easier for data anlysis. For now, just think of tibbles and dataframes as the same thing and do not worry about the difference. 

### Creating Dataframes

We have already worked extensively with dataframes that we have imported through a package or `read.csv`.

```{r}
library(gapminder)
gap <- gapminder
```

We can create a dataframe from scratch using `data.frame()`.  This function takes vectors as input:

```{r}
vec_1 <- 1:3
vec_2 <- c("a", "b", "c")
df <- data.frame(vec_1, vec_2)
```

### The Structure of Dataframes

Under the hood, a dataframe is a list of equal-length vectors. This makes it a 2-dimensional structure, so it shares properties of both the matrix and the list. 

```{r}
vec_1 <- 1:3
vec_2 <- c("a", "b", "c")
df <- data.frame(vec_1, vec_2)

str(df)
```

The `length()` of a dataframe is the length of the underlying list and so is the same as `ncol()`; `nrow()` gives the number of rows. 

```{r}
vec_1 <- 1:3
vec_2 <- c("a", "b", "c")
df <- data.frame(vec_1, vec_2)

# These two are equivalent - number of columns
length(df)
ncol(df)

# Get number of rows
nrow(df)

# Get number of both columns and rows
dim(df)
```

### Naming Dataframes

Dataframes have `colnames()` and `rownames()`. However, since dataframes are really lists (of vectors) under the hood, `names()` and `colnames()` are the same thing. 

```{r}
vec_1 <- 1:3
vec_2 <- c("a", "b", "c")
df <- data.frame(vec_1, vec_2)

# These two are equivalent
names(df)
colnames(df)

# Change the colnames
colnames(df) <- c("Number", "Character")

# Change the rownames
rownames(df) 
rownames(df) <- c("donut", "pickle", "pretzel")
df
```