---
title: "Indexing Dataframes"
author: "PLSC 31101"
date: "Fall 2020"
output: html_document
---

## Indexing Dataframes

> A dataframe is a list that quacks like a matrix.

Remember that dataframes are really lists of vectors (one vector per column). That means that dataframes have both list- and matrix-like behavior.

For example, just as `list$name` selects the `name` element from the list, `df$name` selects the `name` column (vector) from the dataframe:

```{r}
library(gapminder)
gap <- gapminder

head(gap$country)
```

Likewise, we can use square brackets to subset rows and columns:

```{r}
# Row 1, column 3
gap[1, 3]

# Fourth row
gap[4, ]

# First two rows of the columns 1 and 5
gap[c(1,2), c(1, 5)]
```

We can also use subsetting in conjunction with assignment to quickly add a column:

```{r}
names(gap)
gap$new_col <- NA
head(gap)
```

### Challenges

#### Challenge 1. {-}

Create a 3x2 dataframe called `basket`. The first column should contain the names of 3 fruits. The second column should contain the price of those fruits. Now give your dataframe appropriate column and row names.

#### Challenge 2. {-}

Add a third column called `color` that tells what color each fruit is.
