---
title: "10_99-Solutions"
output: html_document
---

## 2. Vectors

#### Challenge 1. {-}

Create a character vector called `fruit` that contains 4 of your favorite fruits. Then evaluate its structure using the commands below:

```{r, eval = FALSE}
# First create your fruit vector 
fruit <- c("raspberries", "mangoes", "lemons", "limes")

# Examine your vector
length(fruit)
class(fruit)
str(fruit)
```

#### Challenge 2. {-}

```{r}
# 1. Create a vector of a sequence of numbers between 1 to 10.
numbers <- c(1:10)

# 2. Coerce that vector into a character vector
numbers_char <- as.character(numbers)

# 3. Add the element "11" to the end of the vector
numbers_char <- c(numbers_char, 11)

# 4. Coerce it back to a numeric vector.
numbers_num <- as.numeric(numbers_char)
```

#### Challenge 3: {-}

Create a vector of the numbers 11 to 20 and multiply it by the original vector from Challenge 2.

```{r}
numbers2 <- c(11:20)
numbers * numbers2
```

## 3. Indexing Vectors

#### Challenge 1. {-}

Subset `country_vector` below to return every value EXCEPT "Canada" and "Brazil".

```{r}
country.vector<-c("Afghanistan", "Canada", "Sierra Leone", "Denmark", "Japan", "Brazil")

# Do it using positive integers
country.vector[c(1, 3, 4, 5)]

# Do it using negative integers
country.vector[-c(2, 6)]

# Do it using a logical vector
country.vector[c(TRUE, FALSE, TRUE, TRUE, TRUE, FALSE)]

# Do it using a conditional statement (and an implicit logical vector)
country.vector[!country.vector %in% c("Canada", "Brazil")]
```

## 5. Indexing Lists

#### Challenge 1. {-}

What are the four basic types of atomic vectors? How does a list differ from an atomic vector?

The four types of atomic vectors are: `integer`, `numeric`, `character`, and `logical`. 
ists are different from vectors because their elements can be of any type, whereas vectors are homogenous.

#### Challenge 2. {-}

Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?

1. R is able to coerce the integer `1` to a character, making it equivalent to the character `"1"`
2. R coerces the logical `FALSE` to the integer `0`, which is greater than `-1`
3. R can coerce the integer `2` to a character, and is not able to compare the two characters with `<`

#### Challenge 3. {-}

Create three vectors and combine them into a list. Assign them names.

```{r}
vec1 <- c("character", "vector")
vec2 <- c(1, 2, 3)
vec3 <- c(TRUE, TRUE, FALSE)

my_list <- list(a = vec1,
                b = vec2,
                c = vec3)
```

#### Challenge 4.  {-}

If `x` is a list, what is the class of `x[1]`?  How about `x[[1]]`?



#### Challenge 5.  {-}

Take a look at the linear model below:

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
summary(mod)
```

Extract the R squared from the model summary.

```{r}
summary(mod)[8]
```

## 7. Dataframes

#### Challenge 1. {-}

Create a 3x2 dataframe called `basket`. The first column should contain the names of 3 fruits. The second column should contain the price of those fruits. Now give your dataframe appropriate column and row names.

```{r}
basket <- data.frame(c("apples", "orange", "bananas"), c(.89, .99, 1.00))

names(basket) <- c("fruit", "price")
```

#### Challenge 2. {-}

Add a third column called `color` that tells what color each fruit is.

```{r}
color = c("red", "orange", "yellow")

basket <- data.frame(basket, color)
```

## 99. Factors

#### Challenge 1.  {-}

What happens to a factor when you modify its levels?
    
```{r, results = "none"}
f1 <- factor(letters)
levels(f1) <- rev(levels(f1))
f1
```

The vector is ordered in reverse, but the factor assignments per letter stay the same.

#### Challenge 2.  {-}

What does this code do? How do `f2` and `f3` differ from `f1`? 

```{r, results = "none"}
f2 <- rev(factor(letters))
f3 <- factor(letters, levels = rev(letters))
```

f2 reverses the vector and assigns the factors in the same order is f1 -- z is assigned to a, etc, by reversing the order of the vector.

f3 similarly assigns z to a, etc, but does so by reversing the order of the factor levels.